\documentclass[draft]{homework}

\usepackage{fixme}


\title{Practical Network Defense - Lab 9}
\subtitle{Vulnerability assessment of ACME co.'s network}
\author{Alessandro Serpi - 1647244}
\date{17 May 2019}


\begin{document}
    \maketitle
    \tableofcontents
    
    
    \pagebreak
    \section{Introduction}
    \fxnote{TODO}
    
    
    \section{Greenbone setup}
    \fxnote{TODO}
    
    
    \section{Rules of engagement}
    \fxnote{TODO}
    
    
    \section{Assessment results and analysis}
    \subsection{Missing cookie attributes in the domain controller}
    \begin{displayquote}
        The application is missing the `httpOnly' attribute.
        \textelp{}
        This allows a cookie to be accessed by JavaScript which could lead to session hijacking attacks.
    \end{displayquote}
    \begin{displayquote}
        The flaw is due to cookie is not using 'secure' attribute, which allows cookie to be passed to the server by the client over non-secure channels (http) and allows attacker to conduct session hijacking attacks.
    \end{displayquote}
    The domain controller does not set `httpOnly' and `secure' cookie attributes, which can allow session hijacking attacks.
    Since this is a Zentyal's shortcoming, we can not implement any mitigation.
    
    \subsection{TCP timestamps}
    \begin{displayquote}
        It was detected that the host implements RFC1323.
        \textelp{}
        A side effect of this feature is that the uptime of the remote host can sometimes be computed.
    \end{displayquote}
    
    An attacker may use TCP timestamps to determine whether security patches requiring a reboot were applied to a host.
    However, disabling timestamps adds security through obscurity, which is no security. 
    In addition, TCP timestamps are used in the PAWS (Protect Against Wrapped Sequence Numbers) mechanism.
    
    Therefore, we decided not to consider the user of this functionality a vulnerability and not to implement any mitigation.
    
    \fxnote{TODO}
    
    
    \section{Mitigations}
    \fxnote{TODO}
    
    
    \section{Final remarks}
    \fxnote{TODO}
\end{document}
